# Preface {.unnumbered}

## Vivax Serology Partnership 

A global partnership between academia, biotechnology industry research and development, non-governmental organisations, policy and implementation stakeholder to address a major gap in the diagnosis of *Plasmodium viavx* infections and accelerate towards elimination. 

For more information see the [VISPA website](https://vispa.online/). 

## Purpose of This Book

This book aims to provide a step-by-step tutorial on how to analyse serology data, with a particular focus on our *Plasmodium vivax* serology test and treat (PvSeroTAT) strategy. 


## SeroTrackR 

`{SeroTrackR}` is a tool to for serology data analysis and visualisation, and for the application of the machine learning algorithms implemented in the `{PvSeroApp}`. The R package is intended to give users more flexibility in assessing serology data and manipulate data visualisations to the user's interest. 
You can download the R package using the following: 

```{r}
#| exec: false
#| eval: false
# Run once to configure your package to use and deploy SeroTrackR
if(!require(devtools)){
  install.packages("devtools") # If not already installed
} 
devtools::install_github("dionnecargy/SeroTrackR")
library(SeroTrackR)
```

Alternatively you can download the package from the `{remotes}` R package: 

```{r}
#| exec: false
#| eval: false
if(!require(remotes)){
  install.packages("remotes") # If not already installed
} 
library(remotes)
remotes::install_github("dionnecargy/SeroTrackR")
```
