---
title: "Bead Counts"
format: html
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
setup <- function() {
  needed <- c("knitr", "rmarkdown", "tidyverse", "kableExtra", "purrr")
  
  lapply(needed, function(pkg) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      library(pkg, character.only = TRUE)
    }
  })
}

setup()
```

## Visualising Bead Counts 

This tutorial focuses on creating plots to visualise the bead counts from your Luminex experimental data. 

### Setup 

```{r setup 1}
#| warning: false
#| message: false
library(SeroTrackR)
library(tidyverse)

your_raw_data <- c(
  system.file("extdata", "example_MAGPIX_plate1.csv", package = "SeroTrackR"),
  system.file("extdata", "example_MAGPIX_plate2.csv", package = "SeroTrackR"), 
  system.file("extdata", "example_MAGPIX_plate3.csv", package = "SeroTrackR")
)
your_plate_layout <- system.file("extdata", "example_platelayout_1.xlsx", package = "SeroTrackR")

sero_data <- readSeroData(
  raw_data = your_raw_data, 
  platform = "magpix",        # default 
  version = "4.2"             # default 
)

plate_list <- readPlateLayout(
  plate_layout = your_plate_layout, 
  sero_data  = sero_data
)

qc_results  <- runQC(
  sero_data = sero_data,       # load in your serological data variable 
  plate_list = plate_list      # load your plate list variable 
)
```

### Plot Bead Counts 

The `plotCounts()` function provides a summary of the bead counts for each plate well are displayed, with blue indicating there are sufficient beads (≥15) or red when there are not enough. If any of the wells are red, they should be double-checked manually and re-run on a new plate if required.

```{r}
plotCounts(qc_results, experiment_name = "experiment1")
```

```{r}
#| exec: false
#| eval: false
getRepeats(qc_results, plate_list)
```

The `getRepeats()` function will inform you whether there are "No repeats necessary" or provide a list of samples to be re-run. In the example data, the beads in plate 2 wells A1 and A2 will need to be repeated. 

```{r}
#| echo: false
getRepeats(qc_results, plate_list) %>%
  kable() %>%
  kable_styling(full_width = TRUE)
```

### Plot Bead Counts per Antigen

This dot plot shows the raw bead counts for each antigen in each plate. This is intended as a quality control check to be able to identify which bead may be an issue.

```{r}
plotBeadCounts(qc_results)
```

## Customisations

For more customisations on how to manipulate your `ggplot2` see the manual [here](https://ggplot2.tidyverse.org/) or the book [@ggplot2]. 