# FAQs

## I have multiple plate layout files. How can I input them?

Use the `getPlateLayout()` function to create a master plate layout file to then input into the other functions in the package!

```{r}
#| exec: false 
#| eval: false
getPlateLayout("your/folder/with/plate/layouts/")
```

Here replace "your/folder/with/plate/layouts/" with the main file that contains your folders. For example, if your folder looks like this:

``` bash
my_R_project/
└── raw_data_files/
    ├── plate_1/
    │   ├── raw_magpix_data_plate1.csv
    │   └── plate_layout_1.xlsx
    ├── plate_2/
    │   ├── raw_magpix_data_plate2.csv
    │   └── plate_layout_2.xlsx
    └── plate_3/
        ├── raw_magpix_data_plate3.csv
        └── plate_layout_3.xlsx
```

you would write:

```{r}
#| exec: false 
#| eval: false
getPlateLayout("raw_data_files/")
```

you could ALSO write:

```{r}
#| exec: false 
#| eval: false
getPlateLayout()
```

OR:

```{r}
#| exec: false
#| eval: false
getPlateLayout(folder_path = c("plate_layout_1.xlsx", "plate_layout_2.xlsx", "plate_layout_3.xlsx"))
```

## I have multiple Luminex data types I'd like to analyse. How can I do this?

If you have, for example, both Bio-Plex and MAGPIX files and would like to analyse them both, then you can do some clever data manipulation as below: 

*Note that in this example, there is one plate layout that contains all files in it. But the same idea can apply to `readPlateLayout()`. 

**Input your Bio-Plex file/s**: 

Using a reproducible example: 

```{r}
#| message: false
#| warning: false
library(SeroTrackR)
library(tidyverse)

bioplex_raw_plates <- c(
  system.file("extdata", "example_BioPlex_plate1.xlsx", package = "SeroTrackR"),
  system.file("extdata", "example_BioPlex_plate2.xlsx", package = "SeroTrackR")
)
all_plate_layout <- system.file("extdata", "example_platelayout_1.xlsx", package = "SeroTrackR")
```

For your data: 

```{r}
#| exec: false
#| eval: false
bioplex_raw_plates <- c(
  "example_BioPlex_plate1.xlsx", 
  "example_BioPlex_plate2.xlsx"
)
all_plate_layout <- "example_platelayout_1.xlsx" 
```

**Input your MAGPIX file/s**: 

Using a reproducible example: 

```{r}
magpix_raw_plate     <- system.file("extdata", "example_MAGPIX_plate3.csv", package = "SeroTrackR")
```

For your data: 

```{r}
#| exec: false
#| eval: false
magpix_raw_plate      <- "example_MAGPIX_plate3.csv"
```

**Read the serological data files in**: 

```{r}
# Serological data
bioplex_sero_data <- readSeroData(
  raw_data = bioplex_raw_plates,
  platform = "bioplex"
)
magpix_sero_data <- readSeroData(
  raw_data = magpix_raw_plate,
  platform = "magpix", 
  version = "4.2"
)
```

**Merge the files together**: 

```{r}
sero_data_merged <- NULL
# data_raw 
sero_data_merged$data_raw <- bioplex_sero_data$data_raw %>% 
  bind_rows(magpix_sero_data$data_raw)

# results 
sero_data_merged$results <- bioplex_sero_data$results %>% 
  bind_rows(magpix_sero_data$results)

# counts 
sero_data_merged$counts <- bioplex_sero_data$counts %>% 
  bind_rows(magpix_sero_data$counts)

# blanks
sero_data_merged$blanks <- bioplex_sero_data$blanks %>% 
  bind_rows(magpix_sero_data$blanks)

# stds
sero_data_merged$stds <- bioplex_sero_data$stds %>% 
  bind_rows(magpix_sero_data$stds)

# run 
sero_data_merged$run <- bioplex_sero_data$run %>% 
  bind_rows(magpix_sero_data$run)
```

**Continue the rest of the pipeline**: 

```{r}
#| exec: false
#| eval: false 
plate_list_all  <- readPlateLayout(
  plate_layout = all_plate_layout, 
  sero_data = sero_data_merged
)

qc_results <- runQC(
  sero_data = sero_data_merged, 
  plate_list = all_plate_layout
)

mfi_to_rau_output <- MFItoRAU(
  sero_data = sero_data_merged,
  plate_list = all_plate_layout, 
  qc_results = qc_results, 
  std_point = 10
)

# etc.. 
```

