---
title: "MFI and RAU plots"
format: html
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
setup <- function() {
  needed <- c("knitr", "rmarkdown", "tidyverse", "kableExtra", "purrr")
  
  lapply(needed, function(pkg) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      library(pkg, character.only = TRUE)
    }
  })
}

setup()
```

## Visualising MFI and RAU Data 

This tutorial focuses on creating boxplots to visualise the Median Fluorescent Intensity (MFI) or converted Relative Antibody Units (RAU) for each of the antigens in the panel. 

The box plots per antigen visualise the min, median, max MFI and IQR values. 

The grey boxplots in the background indicate the MFI values for the dataset used to train the PvSeroApp model.

### Setup 

```{r setup 1}
#| warning: false
#| message: false
library(SeroTrackR)
library(tidyverse)

your_raw_data <- c(
  system.file("extdata", "example_MAGPIX_plate1.csv", package = "SeroTrackR"),
  system.file("extdata", "example_MAGPIX_plate2.csv", package = "SeroTrackR"), 
  system.file("extdata", "example_MAGPIX_plate3.csv", package = "SeroTrackR")
)
your_plate_layout <- system.file("extdata", "example_platelayout_1.xlsx", package = "SeroTrackR")

sero_data <- readSeroData(
  raw_data = your_raw_data, 
  platform = "magpix",        # default 
  version = "4.2"             # default 
)

plate_list <- readPlateLayout(
  plate_layout = your_plate_layout, 
  sero_data  = sero_data
)

qc_results  <- runQC(
  sero_data = sero_data,       # load in your serological data variable 
  plate_list = plate_list      # load your plate list variable 
)

mfi_to_rau_output <- MFItoRAU(
  sero_data = sero_data, 
  plate_list = plate_list,
  qc_results = qc_results, 
  std_point = 10
)
```

### MFI 

We can use the `plotMFI` function inside the SeroTrackR R package to conveniently plot the MFI data from our experiment as well as the grey background boxplots of the WEHI dataset used to train the PvSeroApp model.

```{r}
#| warning: false
#| message: false
plotMFI(mfi_to_rau_output, location = "PNG")
```

You can also plot this using `ggplot2` and customise: 

```{r}
#| warning: false
#| message: false
df_results <- mfi_to_rau_output[[2]]

# relabel antigen names from lab codes to proper antigen names
old_names <- c("EBP", "LF005", "LF010", "LF016", "MSP8", "RBP2b.P87", "PTEX150", "PvCSS")
new_names <- c("PvEBP", "Pv-fam-a", "PvMSP5", "PvMSP1-19",  "PvMSP8", "PvRBP2b", "PvPTEX150", "PvCSS")

name_lookup <- setNames(new_names, old_names)

df_results <- df_results %>%
  select(SampleID, Plate, ends_with("_MFI")) %>%
  rename_with(~str_replace(., "_MFI", ""), ends_with("_MFI")) %>%
  pivot_longer(-c(SampleID, Plate), names_to = "Antigen", values_to = "MFI") %>%
  mutate(
    Plate = factor(Plate, levels = unique(Plate[order(as.numeric(str_extract(Plate, "\\d+")))])), # Reorder by plate number
    MFI = as.numeric(MFI)
  )

df_results %>%
  ggplot(aes(x= Antigen, y = MFI)) +
  geom_boxplot(aes(fill = Antigen)) +
  scale_y_log10(breaks = c(10, 100, 1000, 10000), limits = c(10, 10000), labels = c("10", "100", "1,000", "10,000")) +
  scale_fill_brewer(palette = "Paired", type = "qual") +
  labs(x = "Antigen", y = "Antibody MFI") +
  facet_wrap( ~ Plate) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```

### RAU 

A similar function and code pipeline can be used for RAU. 

We can use the `plotRAU` function inside the SeroTrackR R package to conveniently plot the RAU data from our experiment as well as the grey background boxplots of the WEHI dataset used to train the PvSeroApp model.

```{r}
plotRAU(mfi_to_rau_output, location = "PNG")
```

You can also plot this using `ggplot2` and customise: 

```{r}
df_results <- mfi_to_rau_output[[2]]

# relabel antigen names from lab codes to proper antigen names
old_names <- c("EBP", "LF005", "LF010", "LF016", "MSP8", "RBP2b.P87", "PTEX150", "PvCSS")
new_names <- c("PvEBP", "Pv-fam-a", "PvMSP5", "PvMSP1-19",  "PvMSP8", "PvRBP2b", "PvPTEX150", "PvCSS")

name_lookup <- setNames(new_names, old_names)

df_results <- df_results %>%
  select(SampleID, Plate, ends_with("_Dilution")) %>%
  rename_with(~str_replace(., "_Dilution", ""), ends_with("_Dilution")) %>%
  pivot_longer(-c(SampleID, Plate), names_to = "Antigen", values_to = "RAU") %>%
  mutate(
    Plate = factor(Plate, levels = unique(Plate[order(as.numeric(str_extract(Plate, "\\d+")))])), # Reorder by plate number
    RAU = as.numeric(RAU)
  )

df_results %>%
  ggplot(aes(x= Antigen, y = RAU, fill = Antigen)) +
  geom_boxplot() +
  scale_y_log10(
    breaks = c(1e-5, 1e-4, 1e-3, 1e-2, 0.03),
    labels = c("0.00001", "0.0001", "0.001", "0.01", "0.03")
  ) +
  scale_fill_brewer(palette = "Paired", type = "qual") +
  labs(x = "Antigen", y = "Antibody RAU") +
  facet_wrap( ~ Plate) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Customisations

For more customisations on how to manipulate your `ggplot2` see the manual [here](https://ggplot2.tidyverse.org/) or the book [@ggplot2]. 