---
title: "PvLDH Tutorial"
format: html
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
setup <- function() {
  needed <- c("knitr", "rmarkdown", "tidyverse", "kableExtra", "purrr")
  
  lapply(needed, function(pkg) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      library(pkg, character.only = TRUE)
    }
  })
}

setup()
```

For the PvLDH data analysis, follow the steps below: 

## Using Tutorial Dataset: Load the Data

We will be using the build-in files in the R package for this tutorial, as shown below. 

```{r}
#| warning: false
#| message: false
library(SeroTrackR)
library(tidyverse)

your_raw_LDH_data         <- system.file("extdata", "example_BioPlex_PvLDH_plate1.xlsx", package = "SeroTrackR")
your_LDH_plate_layout     <- system.file("extdata", "example_platelayout_1.xlsx", package = "SeroTrackR")
```

To run your OWN data, follow the code below: 

```{r exec=FALSE, eval=FALSE}
#| exec: false
#| eval: false 
your_raw_LDH_data       <- "PATH/TO/YOUR/FILE/plate1.xlsx"
your_LDH_plate_layout   <- "PATH/TO/YOUR/FILE/plate_layout.xlsx"
```

## Step 1: Read in your data

:::{.callout-note}
### Data preparation 

Please ensure that you have read and prepared your raw Luminex files and plate layout files as per the instructions in the [Before You Begin](intro.qmd) page. 

:::

You can read any Luminex serology data file using the `readSeroData()` function. The Luminex platform can be specified using the `platform` argument, with either `bioplex`, `magpix` or `intelliflex`. 

For the xPONENT software-based exported files (MAGPIX or INTELLIFLEX), the version of the software should be specified as either **"4.2"** or **"4.3"**, with the version **"4.2"** as the default.

```{r}
ldh_sero_data <- readSeroData(
  raw_data = your_raw_LDH_data, 
  platform = "bioplex"
)
```

```{r}
#| exec: false
#| eval: false
ldh_sero_data
```

```{r}
#| results: asis
#| echo: false

iwalk(ldh_sero_data, function(df, nm) {

  cat("$", nm, "\n\n")

  tbl <- kable(df, format = "html") %>%
    kable_styling(full_width = TRUE) %>%
    scroll_box(width = "100%", height = "200px")
  print(tbl)

  cat("\n\n")
})
```

::: {.callout-tip}

### readSeroData Output
This exports a list of data frames: 

1. `raw`: The raw, unedited files: A check that the correct file/s were added.
2. `results`: The processed serological data files: Antigen names compatible with the R package, clear columns for each antigen and their MFI results.
3. `counts`: The bead counts detected from the Luminex instrument per antigen, per sample, for each plate. 
4. `blanks`: The MFI results for the blank wells. 
5. `stds`: The MFI results for the standard curve samples.
6. `run`: The run information from the Luminex machine. 

You can access each data frame, for example the `results` data frame, like this:

```{r}
#| exec: false
#| eval: false
ldh_sero_data$results
```

```{r}
#| echo: false
ldh_sero_data$results %>%
  kable() %>%
  scroll_box(width = "100%", height = "200px")
```

:::

Next, you need to read in the plate layout! This file will allow us to relabel the raw luminex data from "unknowns" to our study Sample ID's. 

```{r}
ldh_plate_list <- readPlateLayout(
  plate_layout = your_LDH_plate_layout, 
  sero_data  = ldh_sero_data
)
```

```{r}
#| exec: false
#| eval: false
ldh_plate_list
```

```{r}
#| results: asis
#| echo: false

iwalk(ldh_plate_list, function(df, nm) {

  cat("$", nm, "\n\n")

  tbl <- kable(df, format = "html") %>%
    kable_styling(full_width = TRUE)
  print(tbl)

  cat("\n\n")
})
```

Great! Now our serological data has been processed and our plate layouts are added into R-compatible data frames ready for analysis!

## Step 2: Quality control checks

Quality control checks can also be run on any serological data file that has been processed using the `readSeroData()` and `readPlateLayout()` functions. The `runQC()` function will: 

1. Process bead count data (assigns a warning if bead counts are less than 15 or above 500 per antigen).
2. Assign Sample ID's to Luminex outputs.

```{r}
ldh_qc_results  <- runQC(
  sero_data = ldh_sero_data,       # load in your serological data variable 
  plate_list = ldh_plate_list      # load your plate list variable 
)
```

::: {.callout-tip}

### runQC Output
This exports a list of data frames: 

1. `processCounts_output``: The bead counts per antigen, per sample, for each plate. Intermediary file for downstream processing. Warning = 0 for pass and = 1 for fail. 
2. `getCounts_output`: The processed bead counts file for each sample, summarising whether a sample needs to be repeated. Intermediary file for downstream processing.
3. `sampleid_output`: Aligning the Luminex anonymised data (SampleID) to the plate layout ID's (Sample). Intermediary file for downstream processing.
4. `getAntigenCounts_output`: The processed bead counts file per antigen per sample. Used for `plotBeadCounts()`. 
5. `getCountsQC_output`: The bead counts and QC results per antigen. **Key file**. 

You can access each data frame, for example the `getCountsQC_output` data frame, like this:

```{r}
#| exec: false
#| eval: false
ldh_qc_results$getCountsQC_output
```

```{r}
#| echo: false
ldh_qc_results$getCountsQC_output %>%
  kable() %>%
  scroll_box(width = "100%", height = "200px")
```

:::

You can also visualise QC results using the `ldh_sero_data` and `ldh_qc_results` variables. 

### Standard Curve Plot

The `plotStds_all()` function plots the standard curve, generated from the antibody data from the standards you indicated in your plate layout (e.g. S1-S10) and Median Fluorescent Intensity (MFI) units are displayed in log10-adjusted scale. 

In the case of the PvSeroTAT multi-antigen panel, the antigens will be displayed and in general your standard curves should look relatively linear (only when the y-axis is on log-adjusted scale). 

```{r}
plotStds_all(ldh_sero_data, experiment_name = "experiment1")
```

For more information on how to interrogate and visualise standard curves, see the [Explore Standard Curves](std_curves.qmd) page. 

### Bead Counts QC Plot

The `plotCounts()` function provides a summary of the bead counts for each plate well are displayed, with blue indicating there are sufficient beads (≥15) or red when there are not enough. If any of the wells are red, they should be double-checked manually and re-run on a new plate if required.

```{r}
plotCounts(ldh_qc_results, experiment_name = "experiment1")
```

```{r}
#| exec: false
#| eval: false
getRepeats(ldh_qc_results, ldh_plate_list)
```

The `getRepeats()` function will inform you whether there are "No repeats necessary" or provide a list of samples to be re-run. In the example data, the beads in plate 2 wells A1 and A2 will need to be repeated. 

```{r}
#| echo: false
getRepeats(ldh_qc_results, ldh_plate_list) %>%
  kable() %>%
  kable_styling(full_width = TRUE)
```

For more information on how to visualise these data, see the [Bead Counts](counts.qmd) page. 

### Blanks QC Plot

The Median Fluorescent Intensity (MFI) units for each antigen is displayed for your blank samples. In general, each blank sample should have ≤50 MFI for each antigen, if they are higher they should be cross-checked manually.

```{r}
plotBlanks(ldh_sero_data, experiment_name = "experiment1")
```

:::{.callout-info}
In the example data, blank samples recorded higher MFI values for LF005 on plate 1 and should be checked to confirm this is expected from the assay.
:::

## Step 3: MFI to RAU conversion 

The automated data processing allows you to convert your Median Fluorescent Intensity (MFI) data into Relative Antibody Units (RAU) by fitting a 5-parameter logistic function to the standard curve on a per-antigen level using the `MFItoRAU_LDH()` function. 

**Key input variables**:
 
- `sero_data`, `plate_list`: Previously defined variables from the `readSeroData()` and `readPlateLayout()` functions. 
- `dilution`: 10-point standard curve dilution: The default is: 1000000, 333333.33, 111111.11, 37037.04, 12345.68, 4115.23, 1371.74, 457.25, 152.42, 50.81 and **does not** need to be specified again unless changing. 

```{r}
#| warning: false
#| message: false
ldh_mfi_to_rau_output <- MFItoRAU_LDH(
  sero_data = ldh_sero_data, 
  plate_list = ldh_plate_list,
  file_path = NULL                      # Defaults to current folder/working directory
)
```

The `mfi_to_rau_output` variable contains the processed MFI and RAU data for each antigen per sample. 

```{r}
#| exec: false
#| eval: false
ldh_mfi_to_rau_output
```

```{r}
#| echo: false
ldh_mfi_to_rau_output %>%
  kable() %>% 
  scroll_box(width = "100%", height = "200px")
```

For more information on how to visualise MFI and RAU data, see the [MFI and RAU plots](mfi_rau.qmd) page. 

## Data Analysis: Run Pipeline 

```{r example code for LDH, eval=FALSE}
runLDHpipeline(
  raw_data = your_raw_LDH_data,
  plate_layout = your_LDH_plate_layout,
  platform = "bioplex",                               # Defaults to "bioplex" but "magpix" or "intelliflex" also works
  file_path = NULL                                    # Defaults to current folder/working directory
)
```
