---
title: "Explore Standard Curves"
format: html
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
setup <- function() {
  needed <- c("knitr", "rmarkdown", "tidyverse", "kableExtra", "purrr")
  
  lapply(needed, function(pkg) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      library(pkg, character.only = TRUE)
    }
  })
}

setup()
```

This tutorial is focussed on how to interrogate standard curves and compare different standard curves from multiple runs.  

## Setup 

```{r setup 1}
#| warning: false
#| message: false
library(SeroTrackR)
library(tidyverse)

your_raw_data <- c(
  system.file("extdata", "example_MAGPIX_plate1.csv", package = "SeroTrackR"),
  system.file("extdata", "example_MAGPIX_plate2.csv", package = "SeroTrackR"), 
  system.file("extdata", "example_MAGPIX_plate3.csv", package = "SeroTrackR")
)
your_plate_layout <- system.file("extdata", "example_platelayout_1.xlsx", package = "SeroTrackR")

sero_data <- readSeroData(
  raw_data = your_raw_data, 
  platform = "magpix",        # default 
  version = "4.2"             # default 
)

plate_list <- readPlateLayout(
  plate_layout = your_plate_layout, 
  sero_data  = sero_data
)
```

## Standard Curves 

The `plotStds()` function plots the standard curve, generated from the antibody data from the standards you indicated in your plate layout (e.g. S1-S10) and Median Fluorescent Intensity (MFI) units are displayed in log10-adjusted scale. 

:::{.callout-important}
For your standard curve, if you have pooled positive control plasma from highly immune adults from: 

- Papua New Guinea (PNG): write `location = "PNG"`
- Ethiopia (ETH): write `location = "ETH"`. 
:::

:::{.callout-note}
Note that the grey dots in the background indicate the standard curve range observed at WEHI for PNG or Ethiopian samples where there are known differences in the standard curve.
:::

```{r}
plotStds(sero_data, location = "PNG", experiment_name = "experiment1")
```

In the case of the PvSeroTAT multi-antigen panel, the antigens will be displayed and in general your standard curves should look relatively linear (only when the y-axis is on log-adjusted scale). 

## Manually Plot Standard Curves

You can also plot the standard curve using `ggplot2` and customise however you wish. 

```{r}
# stratify data
master_file <- sero_data
stds <- master_file$stds

# relabel antigen names from lab codes to proper antigen names
old_names <- c("EBP", "LF005", "LF010", "LF016", "MSP8", "RBP2b.P87", "PTEX150", "PvCSS")
new_names <- c("PvEBP", "Pv-fam-a", "PvMSP5", "PvMSP1-19", "PvMSP8", "PvRBP2b", "PvPTEX150", "PvCSS")

name_lookup <- setNames(new_names, old_names)

stds_1 <- stds %>%
  select(-Location) %>%
  pivot_longer(-c(Sample, Plate), names_to = "Antigen", values_to = "MFI") %>%
  mutate(
    Plate = factor(Plate, levels = unique(Plate[order(as.numeric(str_extract(Plate, "\\d+")))])), # reorder by plate number
    Sample = factor(Sample, levels = unique(Sample[order(as.numeric(str_extract(Sample, "\\d+")))])), #reorder by Standard number
    Antigen = dplyr::recode(Antigen, !!!name_lookup),
    MFI = as.numeric(MFI)
  ) 

stds_1 %>% 
  ggplot() +
  geom_point(aes(x = Sample, y = MFI, color = Plate, group = Plate)) +
  geom_line(data = stds_1, aes(x = Sample, y = MFI, color = Plate, group = Plate)) +
  scale_y_log10(breaks = c(0, 10, 100, 1000, 10000)) +
  labs(
    x = "Standard Curve",
    y = "MFI"
    ) +
  facet_wrap(~Antigen) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Customisations

For more customisations on how to manipulate your `ggplot2` see the manual [here](https://ggplot2.tidyverse.org/) or the book [@ggplot2]. 